Benchmarking Parameterized Quantum Circuit Learning for Quantum Chemical Applications

### 2025-03-04 manuscript/main.tex
manuscript

@@

Like the previously mentioned studies, we wanted to determine whether QML, specifically using PQCs, can offer a quantum advantage (or speedup) on chemical applications.
Herein, we explore two datasets that require the explicit calculation of (RO)CBS-QB3 and CCSD
(RO)CBS-QB3  can be used for X Y and Z as found in the Wood et al pub
CCSD can be used for energies and properties

@@

resilience levels (error mitigation)
none (0) 
1 ``Minimal mitigation costs: Mitigate error associated with readout errors'' ``Twirled Readout Error eXtinction (TREX) measurement twirling''
Medium mitigation cost medium 
2 ``Medium mitigation costs. Typically reduces bias in estimators, but is not guaranteed to be zero-bias.'' ``Level 1 + Zero Noise Extrapolation (ZNE) and gate twirling''

@@

@@

>>>>

Expand on all of these to make a proper case for what we are doing.

@@

Like the previously mentioned studies, we wanted to determine whether QML, specifically using PQCs, can offer a quantum advantage (or speedup) on chemical applications.
Herein, we explore two datasets that require the explicit calculation of (RO)CBS-QB3 and CCSD

BSE49

CCSD can be used for energies and properties

@@

resilience levels (error mitigation)
none (0) 
level 1 readout error mitigation and measurement twirling using Twirled Readout Error eXtinction (TREX) \cite{van_den_berg_model-free_2022}
level 2 level 1 + gate twirling and zero noise extrapolation (ZNE)\cite{kandala_error_2019,li_efficient_2017,temme_error_2017}

@@

BSE49
``This work addresses the aforementioned gap in the literature by constructing a large dataset (4502 datapoints) of computationally predicted BSEs of 49 unique bond types, all of which are determined with a high-level composite theoretical procedure denoted as (RO)CBS-QB3. This approach ensures uniform, high-quality reference data and eliminates the need to collect and verify data gathered from various sources, which may differ substantially in their accuracy. The (RO)CBS-QB3 method is known to produce BDEs of high accuracy. Therefore, it is suitable for developing a database of BSEs that can be used to test and parametrize low-cost computational methods. One particular target application of our dataset is for the training of cost-effective computational approaches like atom-centered potentials (ACPs) or machine learning potentials.''
(RO)CBS-QB3\cite{wood_restricted-open-shell_2006,montgomery_complete_1999,montgomery_complete_2000}

```The structures obtained from the workflow described above were then used for the final step of reference data calculation, using the composite (RO)CBS-QB331,32,33 method. The restricted-open-shell61 CBS-QB3 or ROCBS-QB3 was employed for the open-shell radical fragments, while restricted closed-shell calculations were performed for the closed-shell parent molecules with CBS-QB3. The composite (RO)CBS-QB3 method approximates energies at the complete-basis-set CCSD(T) level, using a series of computationally lower-cost methods including: (i) geometry optimization followed by vibration frequency calculation using the unrestricted-open-shell62 B3LYP/6-311G(2d,d,p) method46,47,48,49,50,51,63, (ii) ROMP2/6-311+G(3d2f,2df,2p) level63,64,65 energy extrapolated to the complete-basis-set limit, (iii) energy calculation at ROMP4(SDQ)/6-31+G(d(f),p) level63,64,66, and (iv) energy calculation at ROCCSD(T)/6-31+G† level63,64,67 (where 6-31+G† is a modified 6-31+G(d) basis set). Note that the final (RO)CBS-QB3 energy includes additional empirical correction terms described in Reference33. Structures were screened to remove any system for which the imaginary frequencies were obtained. The (RO)CBS-QB3 energies for the structures associated with a particular bond breaking reaction were used to obtain the bond separation energies for the dataset.'''

@@

5 qubit/5 ansatz layers: 2*5 + 3 * 5 * 5 = 85 parameters (75 trainable, 5 * 2 features)
A2 $2n$ parameters
HWE-CNOT $3nL$ parameters
2 qubit gates $nL$
Number of parameters (2n+3nL)=(2+3L)*n

$n$, number of qubits and $L$, number of circuit layers

To efficiently run DDCC on IBM Quebec, splitting the data into batches of $\approx 4$ samples with 64 $t_{2}$-amplitudes each.

One iteration would require approximately $N_{\text{samples}} * N_{occ}^{2} * N_{virt}^{2} * N_{\text{shots}} * N_{\text{observables}}$ circuit executions (+ whatever SPSA costs to run per iteration)
